---
title: "Report"
author: "Shruti Pillai"
format: html
editor: visual
---

```{r}
library(tidyverse)
library(ggplot2)
library(kableExtra)
library(lubridate)
```

```{r}
coffee_data <- read_csv("Data/coffee-shop-sales.csv")
```

## Methodology

Before the data is used for analysis, a data cleaning process is performed. All variables are checked for missing values and no missing values are found across the dataset, thus no further action was taken. Next, the class of each variable is verified. The **transaction_date** variable, which contains date information, was stored as a character string and was therefore mutated to a date format to allow for further analysis.

```{r}
#| echo: false
#| output: false
summary(coffee_data)
```

```{r}
#| echo: false
#| output: false
glimpse(coffee_data)
```

```{r}
#| echo: false
#| output: false
coffee_data <- coffee_data |> 
  mutate(transaction_date = as_date(transaction_date, format = "%d/%m/%Y"))
```

Following the goal of determining the best location and best selling products, the relationship between product category, product details, sales quantity, unit price, and store location is explored.

### Summary of data - Sales by location

```{r}
#| echo: false
#| warning: false
#| message: false
#| tbl-cap: "Total sales and quantity sold by store location"
#| label: tbl-saleslocation
coffee_data |> 
  group_by(store_location) |> 
  summarise(sales_quantity = sum(transaction_qty),
            avg_price = round(sum(transaction_qty * unit_price) / sum(transaction_qty), 2),
            total_sales = sum(transaction_qty * unit_price)) |> 
  arrange(desc(total_sales)) |> 
  kable(col.names = c("Store Location", "Quantity Sold", "Average Price", "Total Sales"))
```

@tbl-saleslocation presents the total quantity sold, average product price, and total sales of stores in each location. The store in Hell's Kitchen recorded the highest total sales amount among the three locations, despite having a slightly lower sales quantity compared to the store in Lower Manhattan. Nevertheless, the slightly higher average price contributed to the higher total sales figure.

### Summary of data - Sales by product category

```{r}
#| echo: false
#| warning: false
#| message: false
#| fig-cap: "Sales by product category"
#| label: fig-productsales
coffee_data |> 
  group_by(product_category) |> 
  summarise(sales_quantity = sum(transaction_qty)) |> 
  ggplot(aes(x = reorder(product_category, -sales_quantity), y = sales_quantity, fill = product_category)) +
  geom_col() +
  geom_text(aes(label = sales_quantity), vjust = -0.5, size = 3) +
  scale_fill_brewer(palette = "Paired") +
  scale_y_continuous(labels = scales::comma) +
  labs(x = "Product Category", y = "Quantity Sold") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
      legend.position = "none")
```

@fig-productsales presents the quantity sold by product category. Coffee is the best selling product category, followed by tea. The third best selling category is bakery items; however, the difference between tea and bakery sales is significant, as bakery sold less than half of total tea sold, representing a large gap between the top two best selling categories and the rest.

### Summary of data - Top 5 Best Selling Coffee in Hell's Kitchen

```{r}
#| echo: false
#| warning: false
#| message: false
#| tbl-cap: "Top 5 Best Selling Coffee in Hell's Kitchen"
#| label: tbl-top5coffee
coffee_data |> 
  filter(product_category == "Coffee", store_location == "Hell's Kitchen") |> 
  group_by(product_detail) |> 
  summarise(sales_quantity = sum(transaction_qty),
            total_sales = sum(transaction_qty * unit_price)) |> 
  arrange(desc(sales_quantity)) |> 
  head(5) |> 
  kable(col.names = c("Product Detail", "Quantity Sold", "Total Sales"))
```

As the analysis above establishes that the store in Hell's Kitchen has the highest total sales and coffee is the best selling product category, a further analysis is performed to determine the best selling coffee products for that store. 

@tbl-top5coffee presents the top 5 best selling products within the coffee category for the store located in Hell's Kitchen. **Ouro Brasileiro shot** is the best selling product with the highest sales quantity, followed by **Brazilian Lg** and **Ethiopia Lg**, which are tied at second and third place respectively.
  