---
title: "Identifying High‑Value Locations and Best‑Selling Products"
subtitle: "An Analysis of Coffee Shop Sales Data"
author:
- "Caryn Kesuma"
- "Shruti Pillai"
- "Nattamon Wattanapenpaiboon"
- "Thunchanok Udomittipong"
format: 
  pdf:
    toc: true
    number-sections: true
execute: 
  echo: false
  warning: false
  message: false
bibliography: references.bib
---

```{r}
library(tidyverse)
library(ggplot2)
library(knitr)
library(lubridate)
coffee_data <- read_csv("Data/coffee-shop-sales.csv")
```

# Executive Summary

This report evaluates sales patterns across New York City café locations to identify the strongest site for expansion and the most effective products to prioritise at launch. Although total sales across Hell’s Kitchen, Astoria, and Lower Manhattan differ by less than 3%, Hell’s Kitchen consistently leads in revenue and shows the most stable demand profile. Coffee is the dominant product category, and items that rank highly in both popularity and revenue represent the strongest candidates for a core launch menu. Based on these findings, Hell’s Kitchen is recommended as the optimal location for a new café, supported by a menu centred on high‑demand, high‑revenue beverages.

# Introduction

The central problem addressed in this report is how a new café can choose a location and initial product offering that maximises early sales and customer engagement. The report focuses on two key objectives: identifying the location with the strongest potential for expansion, and determining which product should be prioritised at launch. Together, these findings will support strategic decision‑making for launching a new café with the highest likelihood of early success.

## Dataset Introduction

To tackle this problem, the analysis uses the Coffee Shop Sales dataset from Maven Analytics, which provides detailed transaction‑level information across multiple existing stores. The dataset includes `r nrow(coffee_data)` rows and `r ncol(coffee_data)` columns, representing a large volume of customer purchase activity. Key variables include numeric identifiers such as `transaction_id`, `store_id`, and `product_id`, which uniquely identify each transaction, store, and product. Temporal variables include `transaction_date` (stored as character, which is cleaned to Date object later) and `transaction_time` (stored as a time object), capturing when each purchase occurred. Measures of customer behaviour are represented through `transaction_qty`, a numeric variable indicating the number of items purchased per transaction. Categorical variables (stored as character) such as `store_location`, `product_category`, `product_type`, and `product_detail` describe where the transaction took place and the type of product purchased. Together, these variables provide a rich foundation for analysing sales patterns across locations and product categories.

# Methodology

Before the data is used for analysis, a data cleaning process is performed. All variables are checked for missing values and none are found across the dataset, thus no further action was taken. Next, the class of each variable is verified. The transaction_date variable, which contains date information, was stored as a character string and was therefore converted to a date format to allow for further analysis. Lastly, the data is checked for duplicated rows. As no duplicates are found, all rows are retained for analysis.

```{r}
#| output: false
glimpse(coffee_data)
```

```{r}
#| output: false
coffee_data <- coffee_data |> 
  mutate(transaction_date = as_date(transaction_date, format = "%d/%m/%Y"))
```

```{r}
#| output: false
sum(duplicated(coffee_data))
```

Following the goal of determining the best location and best selling products, the relationship between product category, product details, sales quantity, unit price, and store location is explored.

## Summary of data - Sales by location

```{r}
#| tbl-cap: "Total sales and quantity sold by store location"
#| label: tbl-saleslocation
coffee_data |> 
  group_by(store_location) |> 
  summarise(sales_quantity = sum(transaction_qty),
            avg_price = round(sum(transaction_qty * unit_price) / sum(transaction_qty), 2),
            total_sales = sum(transaction_qty * unit_price),
            total_transactions = n(),
            avg_transaction_value = round(mean(transaction_qty * unit_price, na.rm = TRUE), 2),
            avg_daily_sales = round(total_sales / n_distinct(transaction_date), 2)) |> 
  arrange(desc(total_sales)) |> 
  kable(col.names = c("Store Location", "Quantity Sold", "Average Price", "Total Sales", "Total Transactions", "Average Transaction Values", "Average Daily Sales"))
```

@tbl-saleslocation presents the total quantity sold, average product price, total sales, total transactions, average transaction values, and average daily sales of stores in each location. The store in Hell's Kitchen recorded the highest total sales amount and highest total transaction count among the three locations, despite having a slightly lower sales quantity compared to the store in Lower Manhattan. Nevertheless, its slightly higher average price contributed to the higher total sales figure.

## Summary of data - Sales by product category

```{r}
#| fig-cap: "Sales by product category"
#| label: fig-productsales
coffee_data |> 
  group_by(product_category) |> 
  summarise(sales_quantity = sum(transaction_qty)) |> 
  ggplot(aes(x = reorder(product_category, -sales_quantity), y = sales_quantity, fill = product_category)) +
  geom_col() +
  geom_text(aes(label = sales_quantity), vjust = -0.5, size = 3) +
  scale_fill_brewer(palette = "Paired") +
  scale_y_continuous(labels = scales::comma) +
  labs(x = "Product Category", y = "Quantity Sold") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
      legend.position = "none")
```

@fig-productsales presents the quantity sold by product category. Coffee is the best selling product category, followed by tea. The third best selling category is bakery items; however, the difference between tea and bakery sales is significant, as bakery sold less than half of total tea sold, representing a large gap between the top two best selling categories and the rest.

# Results

## Which Location would be Most Suitable based on Sales?

```{r, include=FALSE}
coffee_sales <- coffee_data %>%
  mutate(sales = transaction_qty * unit_price)

coffee_sales_clean <- coffee_sales %>%
  mutate(month = floor_date(transaction_date, "month"))
```

```{r}
#| output: false
location_summary <- coffee_sales %>%
  group_by(store_location) %>%
  summarise(total_sales = sum(sales),
            total_transactions = n(),
            total_quantity_sold = sum(transaction_qty),
            avg_transaction_value = mean(sales, na.rm = TRUE),
            avg_daily_sales = total_sales / n_distinct(transaction_date)) %>%
  arrange(desc(total_sales)) 
```

### Total Sales by Store Location in NYC

```{r}
#| label: fig-location-sales
ggplot(location_summary,
       aes(x = reorder(store_location, total_sales), y = total_sales)) +
  geom_col(fill = "brown") +
  coord_flip(xlim = NULL, ylim = c(220000, 240000)) +
  labs(title = "Total Sales by Store Location in NYC", x = "Store Location", y = "Total Sales") +
  theme_minimal()
```

@fig-location-sales shows the total sales across the six-month period are similar across all three locations, with **Hell's Kitchen leading, followed by Astoria and Lower Manhattan**.

### Monthly Sales Trends

```{r}
#| label: fig-monthly-sales
#| eval: true
#| echo: false
monthly_sales <- coffee_sales_clean %>%
  group_by(month, store_location) %>%
  summarise(monthly_sales = sum(sales), .groups = "drop")

ggplot(monthly_sales, aes(x = month, y = monthly_sales, colour = store_location)) +
  geom_line(linewidth = 1) +
  geom_point(size = 2) +
  labs(title = "Monthly Sales Trends by Store Location in NYC",
       x = "Month",
       y = "Monthly Sales ($)") +
  theme_minimal()
```

@fig-monthly-sales reveals a **consistent and strong upward trajectory in sales** across all three locations from January to June. All stores dipped slightly in February before recovering and rising gradually to June. This seasonal period likely reflects increased foot traffic in the warmer months, where customers are more likely to leave their homes and visit cafes.

### Peak-Hour Demand Analysis

```{r}
#| label: fig-hourly-sales

coffee_sales_clean <- coffee_sales_clean %>%
  mutate(hour = hour(transaction_time))

hourly_sales <- coffee_sales_clean %>%
  group_by(store_location, hour) %>%
  summarise(sales = sum(sales),
            .groups = "drop")

ggplot(hourly_sales, aes(x = hour, y = sales, colour = store_location)) +
  geom_line(linewidth = 1) +
  labs(title = "Hourly Sales Patterns by Store Location in NYC", 
       x = "Hour of Day",
       y = "Sales ($)") +
  theme_minimal()
```

As shown in @fig-hourly-sales, **all three locations experience the same morning peak**, with sales increasing between 7-10 AM. Hell's Kitchen and Lower Manhattan both reached peaks exceeding \$30,000 in sales during this period, aligning with the high foot traffic from morning commuters.

Lower Manhattan drops to nearly \$0 by 8 PM, and Hell's Kitchen follows a similar pattern, indicating that these locations are correlated with commuter traffic. Meanwhile, Astoria sustains sales throughout the evening.

## Optimal Menu Recommendations

```{r}
#| output: false

coffee_sales_clean |>
  distinct(store_location) |>
  arrange(store_location)

hot_spot <- "Hell's Kitchen"
hells_kitchen_sips <- coffee_sales_clean |>
  filter(store_location == hot_spot)

crowd_faves <- hells_kitchen_sips |>
  group_by(product_detail) |>
  summarise(total_qty = sum(transaction_qty, na.rm = TRUE), .groups = "drop") |>
  arrange(desc(total_qty))

cash_cows <- hells_kitchen_sips |>
  group_by(product_detail) |>
  summarise(total_revenue = sum(sales, na.rm = TRUE), .groups = "drop") |>
  arrange(desc(total_revenue))

crowd_faves
cash_cows

names(coffee_sales_clean)
glimpse(coffee_sales_clean)
str(coffee_sales_clean)
```

```{r}
coffee_sales_clean <- crowd_faves |>
  left_join(cash_cows, by = "product_detail")

top10 <- coffee_sales_clean |>
  arrange(desc(total_revenue)) |>
  slice_head(n = 10) |>
  mutate(product_detail = fct_reorder(product_detail, total_revenue))

```

```{r}
#| label: fig-hk-rev
#| fig-cap: "top 10 drinks by total revenue (hell’s kitchen)"
#| fig-pos: "H"

ggplot(top10, aes(x = total_revenue, y = product_detail)) +
  geom_col(fill = "#2C3E50") +
  scale_x_continuous(labels = scales::dollar_format(prefix = "$")) +
  labs(
    title = "Hell's Kitchen: Top 10 Drinks by Total Revenue",
    x = "Total revenue (USD)",
    y = NULL
  ) +
  theme_minimal(base_size = 12)
```

As shown in @fig-hk-rev, total revenue in Hell’s Kitchen is concentrated among a small set of drinks, with a clear drop-off after the top-ranked items. The highest-earning products include premium coffees and larger-size beverages, suggesting that both higher unit prices and steady demand contribute to revenue. This figure highlights the key drivers of takings and provides an evidence-based starting point for deciding which drinks should be prioritised for menu placement and promotion.

```{r}
#| label: tbl-hk-rev
#| tbl-cap: "top 10 drinks by total revenue (hell’s kitchen)"

tab_rev <- coffee_sales_clean |>
  arrange(desc(total_revenue)) |>
  slice_head(n = 10) |>
  transmute(
    drink = product_detail,
    total_quantity = total_qty,
    total_revenue = total_revenue
  )

knitr::kable(tab_rev, digits = 2)
```

@tbl-hk-rev highlights that the Civet Cat generated the highest total revenue in Hell’s Kitchen (\$7,380), followed by Sustainably Grown Organic Lg (\$7,329.25) and Dark chocolate Lg (\$6,534). Brazilian Lg and Ethiopia Lg also performed strongly (\$5,432 each), suggesting consistent high earners.

```{r}
#| label: tbl-hk-pop
#| tbl-cap: "top 10 drinks by total quantity sold (hell’s kitchen)"

tab_pop <- coffee_sales_clean |>
  arrange(desc(total_qty)) |>
  slice_head(n = 10) |>
  transmute(
    drink = product_detail,
    total_quantity = total_qty,
    total_revenue = total_revenue
  )

knitr::kable(tab_pop, digits = 2)
```

@tbl-hk-pop shows that the Ouro Brasileiro shot was the most popular item (1,854 units), followed by Serenity Green Tea Rg (1,601) and Morning Sunrise Chai Rg (1,589). Brazilian Lg, Ethiopia Lg, and Sustainably Grown Organic Lg appear as both high-volume sellers and high-revenue items, making them strong Core Menu choices.

# Discussion

While Lower Manhattan achieved the highest average transaction value, its overall transaction volume was lower than Hell’s Kitchen, suggesting that revenue may rely more heavily on higher-spending purchases rather than broader customer traffic. Astoria, on the other hand, demonstrated comparable overall sales performance but did not consistently outperform Hell’s Kitchen across the evaluated metrics. As a result, **Hell’s Kitchen represents the strongest balance between customer volume, revenue generation, and consistent demand**.

As shown in @fig-hk-rev and @tbl-hk-rev, the revenue in Hell’s Kitchen is concentrated among a small set of drinks, led by Civet Cat (bringing in a total of \$7,380), Sustainably Grown Organic lg (totalling at \$7,329.25), and Dark Chocolate lg (with \$6,534 in revenue). However, @tbl-hk-pop shows that high revenue does not always correspond to high demand: the Ouro Brasileiro shot is the most popular item (1,854 units), while Civet Cat achieves the top revenue rank despite comparatively low volume (164 units), suggesting a premium price point and “specialty” positioning instead.

Taken together, @tbl-hk-rev and @tbl-hk-pop support a two-tier menu strategy for launch. “Core menu” items should prioritise drinks that perform strongly on both quantity and revenue, because they provide stable demand and reliable contribution to cafe's the overall revenue; for example, Brazilian lg and Ethiopia lg are high-volume sellers (1,552 units each) and also high revenue earners (both coming in at \$5,432 each), while Sustainably Grown Organic lg performs strongly on both metrics (1,543 units sold with a revenue of \$7,329.25). In contrast, drinks that rank highly in @tbl-hk-rev but not in @tbl-hk-pop (most notably Civet Cat) are best treated as “premium” options to increase the average transaction value through targeted promotion (e.g., staff recommendations or featured placement), while core items are left to anchor consistent day-to-day sales.

# Conclusion

Tables @tbl-hk-rev and @tbl-hk-pop together identify the best menu candidates by combining profitability and demand. Items near the top of both tables are strong “core menu” choices because they sell frequently and generate high total revenue, making them reliable for day-to-day operations. Drinks that rank highly in @tbl-hk-rev but not @tbl-hk-pop likely achieve revenue through a higher price point; these are good premium offerings to keep for margin and upselling, but they may require targeted promotion and careful stock planning. Conversely, drinks that are very popular in @tbl-hk-pop but lower in @tbl-hk-rev are dependable volume sellers and can be used as anchors for bundles or add-ons. Overall, a balanced menu should prioritise overlapping winners, then supplement with a small set of premium specials.

Furthermore, our analysis also examined the sales performance of Hell's Kitchen, Astoria, and Lower Manhattan, across total sales, transaction behaviour, monthly trends, and hourly patterns. All three locations performed similarly in aggregate, with total sales differing by less than 3% over the six-month period. Monthly trends were consistent across locations, with a dip in February followed by sustained growth through to June, indicating that seasonal demand, rather than location-specific factors, is the dominant driver of revenue fluctuations.

Taking all findings together, **Hell's Kitchen is recommended as the strongest candidate for a new café**. Its more evenly distributed demand and competitive overall sales provide a more stable and sustainable commercial foundation, which are vital for a new café seeking to establish a consistent customer base from the start.

# Recommendations

-   The new cafe should be established in Hell's Kitchen.
-   Business operations should prioritise maintaining stable day-to-day customer demand rather than focusing only on high-calue transactions.
-   Profits can be optimised by combining profitability and demand of drinks when choosing menu options.
-   Drinks for the menu should be chosen from those which can be identified in @tbl-hk-rev and @tbl-hk-pop

# References

Free Sample Dataset Download - Coffee Shop Sales - Maven Analytics \| Build Data Skills, Faster. <https://mavenanalytics.io/data-playground/coffee-shop-sales.> Accessed 21 May 2026.

::: hidden
@tidyverse @ggplot2 @knitr @lubridate
:::
