---
title: "Identifying High‑Value Locations and Best‑Selling Products"
subtitle: "An Analysis of Coffee Shop Sales Data"
author:
- "Caryn Kesuma"
- "Shruti Pillai"
- "Nattamon Wattanapenpaiboon"
- "Thunchanok Udomittipong"
format: 
  pdf:
    toc: true
    number-sections: true
---

```{r}
library(tidyverse)
library(ggplot2)
library(kableExtra)
library(lubridate)
coffee_data <- read_csv("Data/coffee-shop-sales.csv")
```

# Results

```{r, include=FALSE}
coffee_data <- read_csv("Data/coffee-shop-sales.csv")

# prep data
coffee_data <- coffee_data %>%
  mutate(transaction_date = dmy(transaction_date))

coffee_sales <- coffee_data %>%
  mutate(sales = transaction_qty * unit_price)

coffee_sales_clean <- coffee_sales %>%
  mutate(month = floor_date(transaction_date, "month"))
```

## Which Location would be Most Suitable based on Sales?

### Overall Store Performance

```{r}
#| label: tbl-location-summary
#| tbl-cap: "Summary statistics by store location in New York City."

location_summary <- coffee_sales %>%
  group_by(store_location) %>%
  summarise(total_sales = sum(sales),
            total_transactions = n(),
            total_quantity_sold = sum(transaction_qty),
            avg_transaction_value = mean(sales, na.rm = TRUE),
            avg_daily_sales = total_sales / n_distinct(transaction_date)) %>%
  arrange(desc(total_sales))

location_summary
```

@tbl-location-summary presents the overall performance metrics across all three NYC locations for the six-month period.

1.  *Hell's Kitchen* led on total sales (\$236,511), total transactions (50,735), and average daily sales (\$1,306.69), suggesting it attracts the highest volume of customers consistently.
2.  *Astoria* came in second, recording \$232,244 in total sales and an average daily sales figure of \$1,283.12.
3.  *Lower Manhattan* ranked third in total sales (\$230,057) and had the fewest transactions (47,782), but has the highest average transaction value at \$4.81 (which is interestingly higher than that of Hell's Kitchen (\$4.66) and Astoria (\$4.59)).

Across all three locations, the differences in total sales are narrow - spanning just \~\$6,500 from top to bottom - indicating a broadly competitive and fairly matched market.

### Total Sales by Store Location in NYC

```{r}
#| label: fig-location-sales
ggplot(location_summary,
       aes(x = reorder(store_location, total_sales), y = total_sales)) +
  geom_col(fill = "brown") +
  coord_flip(xlim = NULL, ylim = c(220000, 240000)) +
  labs(title = "Total Sales by Store Location in NYC", x = "Store Location", y = "Total Sales") +
  theme_minimal()
```

@fig-location-sales shows the total sales across the six-month period are similar across all three locations, with **Hell's Kitchen leading, followed by Astoria and Lower Manhattan**.

### Monthly Sales Trends

```{r}
#| label: fig-monthly-sales

coffee_sales_clean <- coffee_sales %>%
  mutate(transaction_date = dmy(transaction_date),
         month = floor_date(transaction_date, "month"))

monthly_sales <- coffee_sales_clean %>%
  group_by(month, store_location) %>%
  summarise(monthly_sales = sum(sales), .groups = "drop")

ggplot(monthly_sales, aes(x = month, y = monthly_sales, colour = store_location)) +
  geom_line(linewidth = 1) +
  geom_point(size = 2) +
  labs(title = "Monthly Sales Trends by Store Location in NYC",
       x = "Month",
       y = "Monthly Sales ($)") +
  theme_minimal()
```

@fig-monthly-sales reveals a **consistent and strong upward trajectory in sales** across all three locations from January to June. All stores dipped slightly in February before recovering and rising gradually to June. This seasonal period likely reflects increased foot traffic in the warmer months, where customers are more likely to leave their homes and visit coffee shops.

By June, Hell's Kitchen pulled ahead with the highest monthly sales (\~\$57,000), followed by Astoria (\~\$55,500) and Lower Manhattan (\~\$54,500). The convergence of all three lines through March and April indicates similar demand between all three locations during these months, while the slight divergence in May to June suggests **Hell's Kitchen has a modest but growing edge heading into summer**.

## Conclusion

This analysis examined the sales performance of Hell's Kitchen, Astoria, and Lower Manhattan, across total sales, transaction behaviour, monthly trends, and hourly patterns.

Taking all findings together, **Hell's Kitchen is recommended as the strongest candidate location for a new café**. Its more evenly distributed demand and competitive overall sales provide a more stable and sustainable commercial foundation, which are vital for a new café seeking to establish a consistent customer base from the start.
