---
title: "Identifying High‑Value Locations and Best‑Selling Products"
subtitle: "An Analysis of Coffee Shop Sales Data"
author:
- "Caryn Kesuma"
- "Shruti Pillai"
- "Nattamon Wattanapenpaiboon"
- "Thunchanok Udomittipong"
format: 
  pdf:
    toc: true
    number-sections: true
editor: visual
execute: 
  echo: false
  warning: false
---

```{r}
library(tidyverse)
library(ggplot2)
library(kableExtra)
library(lubridate)
coffee_data <- read_csv("Data/coffee-shop-sales.csv")
```


## Executive Summary

## Introduction
The central problem addressed in this report is how a new café can choose a location and initial product offering that maximises early sales and customer engagement. The report focuses on two key objectives: identifying the location with the strongest potential for expansion, and determining which product should be prioritised at launch. Together, these findings will support strategic decision‑making for launching a new café with the highest likelihood of early success.

## Dataset Introduction
To tackle this problem, the analysis uses the Coffee Shop Sales dataset from Maven Analytics, which provides detailed transaction‑level information across multiple existing stores. The dataset includes `r nrow(coffee_data)` rows and `r ncol(coffee_data)` columns, representing a large volume of customer purchase activity. Key variables include numeric identifiers such as `transaction_id`, `store_id`, and `product_id`, which uniquely identify each transaction, store, and product. Temporal variables include `transaction_date` (stored as character) and `transaction_time` (stored as a time object), capturing when each purchase occurred. Measures of customer behaviour are represented through `transaction_qty`, a numeric variable indicating the number of items purchased per transaction. Categorical variables (stored as character) such as `store_location`, `product_category`, `product_type`, and `product_detail` describe where the transaction took place and the type of product purchased. Together, these variables provide a rich foundation for analysing sales patterns across locations and product categories.

## Methodology

Before the data is used for analysis, a data cleaning process is performed. All variables are checked for missing values and no missing values are found across the dataset, thus no further action was taken. Next, the class of each variable is verified. The **transaction_date** variable, which contains date information, was stored as a character string and was therefore mutated to a date format to allow for further analysis.

```{r}
#| echo: false
#| output: false
summary(coffee_data)
```

```{r}
#| echo: false
#| output: false
glimpse(coffee_data)
```

```{r}
#| echo: false
#| output: false
coffee_data <- coffee_data |> 
  mutate(transaction_date = as_date(transaction_date, format = "%d/%m/%Y"))
```

Following the goal of determining the best location and best selling products, the relationship between product category, product details, sales quantity, unit price, and store location is explored.

### Summary of data - Sales by location

```{r}
#| echo: false
#| warning: false
#| message: false
#| tbl-cap: "Total sales and quantity sold by store location"
#| label: tbl-saleslocation
coffee_data |> 
  group_by(store_location) |> 
  summarise(sales_quantity = sum(transaction_qty),
            avg_price = round(sum(transaction_qty * unit_price) / sum(transaction_qty), 2),
            total_sales = sum(transaction_qty * unit_price)) |> 
  arrange(desc(total_sales)) |> 
  kable(col.names = c("Store Location", "Quantity Sold", "Average Price", "Total Sales"))
```

@tbl-saleslocation presents the total quantity sold, average product price, and total sales of stores in each location. The store in Hell's Kitchen recorded the highest total sales amount among the three locations, despite having a slightly lower sales quantity compared to the store in Lower Manhattan. Nevertheless, the slightly higher average price contributed to the higher total sales figure.

### Summary of data - Sales by product category

```{r}
#| echo: false
#| warning: false
#| message: false
#| fig-cap: "Sales by product category"
#| label: fig-productsales
coffee_data |> 
  group_by(product_category) |> 
  summarise(sales_quantity = sum(transaction_qty)) |> 
  ggplot(aes(x = reorder(product_category, -sales_quantity), y = sales_quantity, fill = product_category)) +
  geom_col() +
  geom_text(aes(label = sales_quantity), vjust = -0.5, size = 3) +
  scale_fill_brewer(palette = "Paired") +
  scale_y_continuous(labels = scales::comma) +
  labs(x = "Product Category", y = "Quantity Sold") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
      legend.position = "none")
```

@fig-productsales presents the quantity sold by product category. Coffee is the best selling product category, followed by tea. The third best selling category is bakery items; however, the difference between tea and bakery sales is significant, as bakery sold less than half of total tea sold, representing a large gap between the top two best selling categories and the rest.

