---
title: "Report"
author: "Shruti Pillai"
format: html
editor: visual
---
```{r, include = FALSE}
library(tidyverse)
library(ggplot2)
library(knitr)
library(lubridate)
coffee_data <- read_csv("Data/coffee-shop-sales.csv")
```

```{r, include = FALSE}
coffee_data <- read_csv("Data/coffee-shop-sales.csv")

# prep data
coffee_data <- coffee_data %>%
  mutate(transaction_date = dmy(transaction_date))

coffee_sales <- coffee_data %>%
  mutate(sales = transaction_qty * unit_price)

coffee_sales_clean <- coffee_sales %>%
  mutate(month = floor_date(transaction_date, "month"))
```

# Analysis

## Optimal Menu Recommendations

```{r, echo = FALSE}
#| output: false

coffee_sales_clean |>
  distinct(store_location) |>
  arrange(store_location)

hot_spot <- "Hell's Kitchen"
hells_kitchen_sips <- coffee_sales_clean |>
  filter(store_location == hot_spot)

crowd_faves <- hells_kitchen_sips |>
  group_by(product_detail) |>
  summarise(total_qty = sum(transaction_qty, na.rm = TRUE), .groups = "drop") |>
  arrange(desc(total_qty))

cash_cows <- hells_kitchen_sips |>
  group_by(product_detail) |>
  summarise(total_revenue = sum(sales, na.rm = TRUE), .groups = "drop") |>
  arrange(desc(total_revenue))

crowd_faves
cash_cows

names(coffee_sales_clean)
glimpse(coffee_sales_clean)
str(coffee_sales_clean)
```

```{r, echo = FALSE}

coffee_sales_clean <- crowd_faves |>
  left_join(cash_cows, by = "product_detail")

top10 <- coffee_sales_clean |>
  arrange(desc(total_revenue)) |>
  slice_head(n = 10) |>
  mutate(product_detail = fct_reorder(product_detail, total_revenue))

```

```{r, echo = FALSE}
#| label: fig-hk-rev
#| fig-cap: "top 10 drinks by total revenue (hell’s kitchen)"
#| fig-pos: "H"

ggplot(top10, aes(x = total_revenue, y = product_detail)) +
  geom_col(fill = "#2C3E50") +
  scale_x_continuous(labels = scales::dollar_format(prefix = "$")) +
  labs(
    title = "Hell's Kitchen: Top 10 Drinks by Total Revenue",
    x = "Total revenue (USD)",
    y = NULL
  ) +
  theme_minimal(base_size = 12)
```

As shown in @fig-hk-rev, total revenue in Hell’s Kitchen is concentrated among a small set of drinks, with a clear drop-off after the top-ranked items. The highest-earning products include premium coffees and larger-size beverages, suggesting that both higher unit prices and steady demand contribute to revenue. This figure highlights the key drivers of takings and provides an evidence-based starting point for deciding which drinks should be prioritised for menu placement and promotion.

```{r, echo = FALSE}
#| label: tbl-hk-rev
#| tbl-cap: "top 10 drinks by total revenue (hell’s kitchen)"

tab_rev <- coffee_sales_clean |>
  arrange(desc(total_revenue)) |>
  slice_head(n = 10) |>
  transmute(
    drink = product_detail,
    total_quantity = total_qty,
    total_revenue = total_revenue
  )

knitr::kable(tab_rev, digits = 2)
```

@tbl-hk-rev highlights that the Civet Cat generated the highest total revenue in Hell’s Kitchen (\$7,380), followed by Sustainably Grown Organic Lg (\$7,329.25) and Dark chocolate Lg (\$6,534). Brazilian Lg and Ethiopia Lg also performed strongly (\$5,432 each), suggesting consistent high earners.

```{r, echo = FALSE}
#| label: tbl-hk-pop
#| tbl-cap: "top 10 drinks by total quantity sold (hell’s kitchen)"

tab_pop <- coffee_sales_clean |>
  arrange(desc(total_qty)) |>
  slice_head(n = 10) |>
  transmute(
    drink = product_detail,
    total_quantity = total_qty,
    total_revenue = total_revenue
  )

knitr::kable(tab_pop, digits = 2)
```

@tbl-hk-pop shows that the Ouro Brasileiro shot was the most popular item (1,854 units), followed by Serenity Green Tea Rg (1,601) and Morning Sunrise Chai Rg (1,589). Brazilian Lg, Ethiopia Lg, and Sustainably Grown Organic Lg appear as both high-volume sellers and high-revenue items, making them strong Core Menu choices.

# Discussion

As shown in @fig-hk-rev and @tbl-hk-rev, the revenue in Hell’s Kitchen is concentrated among a small set of drinks, led by Civet Cat (bringing in a total of \$7,380), Sustainably Grown Organic lg (totalling at \$7,329.25), and Dark Chocolate lg (with \$6,534 in revenue). However, @tbl-hk-pop shows that high revenue does not always correspond to high demand: the Ouro Brasileiro shot is the most popular item (1,854 units), while Civet Cat achieves the top revenue rank despite comparatively low volume (164 units), suggesting a premium price point and “specialty” positioning instead.

Taken together, @tbl-hk-rev and @tbl-hk-pop support a two-tier menu strategy for launch. “Core menu” items should prioritise drinks that perform strongly on both quantity and revenue, because they provide stable demand and reliable contribution to cafe's the overall revenue; for example, Brazilian lg and Ethiopia lg are high-volume sellers (1,552 units each) and also high revenue earners (both coming in at \$5,432 each), while Sustainably Grown Organic lg performs strongly on both metrics (1,543 units sold with a revenue of $7,329.25). In contrast, drinks that rank highly in @tbl-hk-rev but not in @tbl-hk-pop (most notably Civet Cat) are best treated as “premium” options to increase the average transaction value through targeted promotion (e.g., staff recommendations or featured placement), while core items are left to anchor consistent day-to-day sales.

# Conclusion

Tables @tbl-hk-rev and @tbl-hk-pop together identify the best menu candidates by combining profitability and demand. Items near the top of both tables are strong “core menu” choices because they sell frequently and generate high total revenue, making them reliable for day-to-day operations. Drinks that rank highly in @tbl-hk-rev but not @tbl-hk-pop likely achieve revenue through a higher price point; these are good premium offerings to keep for margin and upselling, but they may require targeted promotion and careful stock planning. Conversely, drinks that are very popular in @tbl-hk-pop but lower in @tbl-hk-rev are dependable volume sellers and can be used as anchors for bundles or add-ons. Overall, a balanced menu should prioritise overlapping winners, then supplement with a small set of premium specials.

# Recommendations

- The new cafe should be opened in Hell's Kitchen.
- Profits can be optimised by combining profitability and demand of drinks when choosing menu options.
- Drinks for the menu should be chosen from those which can be identified in @tbl-hk-rev and @tbl-hk-pop