---
title: "Identifying High‑Value Locations and Best‑Selling Products"
subtitle: "An Analysis of Coffee Shop Sales Data"
author:
- "Caryn Kesuma"
- "Shruti Pillai"
- "Nattamon Wattanapenpaiboon"
- "Thunchanok Udomittipong"
format: revealjs
editor: visual
execute: 
  echo: false
  warning: false
---

```{r}
library(tidyverse)
coffee_data <- read_csv("Data/coffee-shop-sales.csv")
```

## Introduction

How can a new café choose a location and initial product offering that maximises early sales and customer engagement?

**Objectives:**

\- identifying the location with the strongest potential for expansion

\- determining which product should be prioritised at launch

Together, these findings will support strategic decision‑making for launching a new café with the highest likelihood of early success.

## Dataset Introduction

```{r}
data_dict <- tibble(
  variable = names(coffee_data),
  type = sapply(coffee_data, class),
  description = c("Unique sequential ID representing an individual transaction",
                         "Date of the transaction (MM/DD/YY)", 
                         "Timestamp of the transaction (HH:MM:SS)",
                         "Quantity of items sold",
                         "Unique ID of the coffee shop where the transaction took place",
                         "Location of the coffee shop where the transaction took place",
                         "Unique ID of the product sold",
                         "Retail price of the product sold",
                         "Description of the product category",
                         "Description of the product type",
                         "Description of the product detail"
                         ))


library(knitr)
kable(data_dict)
```
